### 2021-03-22 ch/6-discussion.tex
Submitted version

Finally, an important issue is the communication between human and machine.
The methods presented in this thesis have all strived to make it easy for human operators to understand what is going on:
the method in \chapref{ch:paper1} is configured with simple statistics;
the method in \chapref{ch:paper2} produces invariant images that can be inspected;
and the method in \chapref{ch:paper3} has the same capability
(though it is not part of the core functionality).
However, this is only a beginning.
It is still not easy to notice that a given dataset has problems,
and machines on their own cannot notice the problem,
because they do not (yet) have a complete understanding about what it is that humans care about.
Human preferences are incredibly complex and hard to predict for today's algorithms.
Thus, it is important that machines get feedback early and often,
to keep them aligned with human goals.
(The aforementioned \citet{stiennon2020learning} is arguably an example of this.)
Applied to the problem of dataset bias,
this could mean visualising correlations in the dataset,
or routinely producing invariant representations to show what the network thinks it is meant to learn.

>>>>

Finally, an important issue is the communication between human and machine.
The methods presented in this thesis have all strived to make it easy for human operators to understand what is going on:
the method in \chapref{ch:paper1} is configured with simple statistics;
the method in \chapref{ch:paper2} produces invariant images that can be inspected;
and the method in \chapref{ch:paper3} has the same capability
(though it is not part of the core functionality).
However, this is only a beginning.
It is still not easy to notice that a given dataset has problems,
and machines on their own cannot notice the problem,
because they do not (yet) have a complete understanding about what it is that humans care about;
human preferences are incredibly complex \citep{yudkowsky2011complex} and hard to predict for today's algorithms.
Thus, it is important that machines get feedback early and often,
to keep them aligned with human goals.
(The aforementioned \citet{stiennon2020learning} is arguably an example of this.)
Applied to the problem of dataset bias,
this could mean visualising correlations in the dataset,
or routinely producing invariant representations to show what the network thinks it is meant to learn.

### 2021-03-22 paper3/paper3.tex
Submitted version

\paragraph{Perfect bag.}
We call a sampled set for which $y\perp s$ holds, a perfect bag \citep{chouldechova2017fair,kleinberg2016inherent}.
Such sets are also very desirable when training algorithmically fair classifiers
(see the \emph{sampling} method in \citet{kamiran2012data}).
%
%If we have access to perfect bags, we could equalise true positive/negative rates (\eqref{eq:TPR}) and also equalise positive/negative predicted values (\eqref{eq:PPV}) for all demographic groups.
%This can be shown by using the sum and product rule of conditional probabilities, e.g. \citet{KanRotZia19}.
%We consider a binary-valued subgroup label, $s\sp{\prime}$ versus $s\sp{\prime\prime}$. 
%For $s\sp{\prime}$, we can compute: 
%$P(y=1|\hat{y}=1,s\sp{\prime})=$
%%\begin{align}
%$\nicefrac{P(\hat{y}=1|y=1,s\sp{\prime})P(y=1|s\sp{\prime})}{\left(A + B\right)}$, with $A = P(\hat{y}=1|y=1,s\sp{\prime})P(y=1|s\sp{\prime})$ and $B = P(\hat{y}=1|y=0,s\sp{\prime})(1-P(y=1|s\sp{\prime}))$.
%%\end{align}
%We can do similarly for $s\sp{\prime\prime}$.
%The conditional probability on the left hand side is 
%a positive predicted value, and this quantity can be expressed in terms of true positive/negative rates and the base (prior) rate, shown on the right hand side. 
%If we have a perfect bag ($y\perp s$ holds, which means equal base rates $P(y=1|s\sp{\prime}) = P(y=1|s\sp{\prime\prime})$), an equality in the true positive/negative rates will give us an equality in the positive/negative predicted values.
%Similarly, with a perfect bag, we can equalise true positive/negative rates (eq. \ref{eq:TPR}) and also positive rates (eq. \ref{eq:AR}) for all demographic groups.
%From the sum probability rule, we have:
%$
%P(\hat{y}=1|s\sp{\prime})
%= P(\hat{y}=1|y=1,s\sp{\prime})P(y=1|s\sp{\prime}) + P(\hat{y}=1|y=0,s\sp{\prime})(1-P(y=1|s\sp{\prime}))
%$ for $s\sp{\prime}$ value, and accordingly for $s\sp{\prime\prime}$ value.
%Here, a positive rate on the left hand side is related to true positive/negative rates and the base (prior) rate as shown on the right hand side. 
%
Ideally, we would like to sample our deployment dataset as perfect bags.
However, the deployment set is unlabelled and is unlikely to be perfect in practice. 
Instead, we pursue learning under zero-label systematic bias as learning disentangled representations with a collection of \emph{approximately} perfect bags. We show that the disentangling procedure is robust enough to work with this relaxation, but that performance scales with how well the deployment set is balanced.

>>>>

\paragraph{Perfect bag.}
We call a sampled set for which $y\perp s$ holds, a perfect bag \citep{chouldechova2017fair,kleinberg2016inherent}.
Such sets are also very desirable when training algorithmically fair classifiers
(see the \emph{sampling} method in \citet{kamiran2012data}).
%
%If we have access to perfect bags, we could equalise true positive/negative rates (\eqref{eq:TPR}) and also equalise positive/negative predicted values (\eqref{eq:PPV}) for all demographic groups.
%This can be shown by using the sum and product rule of conditional probabilities, e.g. \citet{KanRotZia19}.
%We consider a binary-valued subgroup label, $s\sp{\prime}$ versus $s\sp{\prime\prime}$. 
%For $s\sp{\prime}$, we can compute: 
%$P(y=1|\hat{y}=1,s\sp{\prime})=$
%%\begin{align}
%$\nicefrac{P(\hat{y}=1|y=1,s\sp{\prime})P(y=1|s\sp{\prime})}{\left(A + B\right)}$, with $A = P(\hat{y}=1|y=1,s\sp{\prime})P(y=1|s\sp{\prime})$ and $B = P(\hat{y}=1|y=0,s\sp{\prime})(1-P(y=1|s\sp{\prime}))$.
%%\end{align}
%We can do similarly for $s\sp{\prime\prime}$.
%The conditional probability on the left hand side is 
%a positive predicted value, and this quantity can be expressed in terms of true positive/negative rates and the base (prior) rate, shown on the right hand side. 
%If we have a perfect bag ($y\perp s$ holds, which means equal base rates $P(y=1|s\sp{\prime}) = P(y=1|s\sp{\prime\prime})$), an equality in the true positive/negative rates will give us an equality in the positive/negative predicted values.
%Similarly, with a perfect bag, we can equalise true positive/negative rates (eq. \ref{eq:TPR}) and also positive rates (eq. \ref{eq:AR}) for all demographic groups.
%From the sum probability rule, we have:
%$
%P(\hat{y}=1|s\sp{\prime})
%= P(\hat{y}=1|y=1,s\sp{\prime})P(y=1|s\sp{\prime}) + P(\hat{y}=1|y=0,s\sp{\prime})(1-P(y=1|s\sp{\prime}))
%$ for $s\sp{\prime}$ value, and accordingly for $s\sp{\prime\prime}$ value.
%Here, a positive rate on the left hand side is related to true positive/negative rates and the base (prior) rate as shown on the right hand side. 
%
Ideally, we would like to sample our deployment dataset as perfect bags.
However, the deployment set is unlabelled and is unlikely to be perfect in practice. 
Instead, we pursue learning under zero-label systematic bias as learning disentangled representations with a collection of \emph{approximately} perfect bags (produced with clustering techniques, see section~\ref{sec:implementation}).
We show that the disentangling procedure is robust enough to work with this relaxation, but that performance scales with how well the deployment set is balanced.
